---
title: 'Challenge #2'
author: "Jyhreh Johnson"
date: "4/19/2022"
output: 
  html_document: 
    highlight: zenburn
    theme: readable
---

## Challenge 2
```{r}
library(tidyverse)
library(tidyr)
library(mosaic)
library(car)
library(broom)
library(ggplot2)
library(infer)
library(dplyr)
library(skimr)
library(ggpubr)
```

## Preliminaries

• Using the {tidyverse} read_csv() function, load the “AVONETdataset1.csv” dataset from this URL as a “tibble” named d.
```{r}
f2 <- "https://raw.githubusercontent.com/difiore/ada-2022-datasets/main/AVONETdataset1.csv"
d2 <- read_csv(f2, col_names = TRUE)
head(d2)
```

• Winnow the dataset to include only the following variables: Species1, Family1, Order1, Beak.Length_Culmen, Beak.Width, Beak.Depth, Tarsus.Length, Wing.Length, Tail.Length, Mass, Habitat, Migration, Trophic.Level, Trophic.Niche, Min.Latitude, Max.Latitude, Centroid.Latitude, Primary.Lifestyle, ** and Range.Size
```{r}
d2 <- d2 %>%
  select(Species1, 
         Family1, 
         Order1, 
         Beak.Length_Culmen, 
         Beak.Width, Beak.Depth, 
         Tarsus.Length, 
         Wing.Length, 
         Tail.Length, 
         Mass, 
         Habitat, 
         Migration, 
         Trophic.Level, 
         Trophic.Niche, 
         Primary.Lifestyle, 
         Min.Latitude, 
         Max.Latitude, 
         Centroid.Latitude, 
         Range.Size)
```
## Step 1

• Create the following two new variables and add them to AVONET dataset:
Relative beak length, which you should calculate as the residual of log(Beak.Length_Culmen) on log(Mass).
```{r}
d2 <- d2 %>% mutate(logBL = log(Beak.Length_Culmen), logMass = log(Mass))
blm <- lm(data = d2, logBL ~ logMass) #linear model
summary(blm) #summarize
blm$residuals #get residuals
head(blm$residuals)
```

• Relative tarsus length, which you should calculate as the residual of log(Tarsus.Length) on log(Mass).
```{r}
d2 <- d2 %>% mutate(logTL = log(Tarsus.Length), logMass = log(Mass))
tlm <- lm(data = d2, logTL ~ logMass) #linear modl
summary(tlm) #summarize
tlm$residuals #get residuals
head(tlm$residuals)
```
## Step 2

• Make a boxplot or violin plot of your new relative tarsus length variable in relation to Primary.Lifestyle 
```{r}
#create new fields for relative beak length = to its residuals and relative tarsus length = to its residuals
d2 <- d2 %>%
  mutate(rel_beak_len = blm$residuals) %>%
  mutate(rel_tar_len = tlm$residuals)
```

```{r}
#Create violin plot
t_lifestyle <- ggplot(d2, aes(x = rel_tar_len, y = Primary.Lifestyle)) + geom_violin()
t_lifestyle
```
our new relative beak length variable in relation to Trophic.Niche
```{r}
#Create boxplot
bk_lifestyle <- ggplot(d2, aes(x = rel_beak_len, y = Trophic.Niche)) + geom_boxplot()
bk_lifestyle
```
## Step 3

• Run an ANOVA analyses to look at the association between geographic range size and the variable Migration. You should first drop those observations for which Migration is not scored and also look at the distribution of the variable Range.Size to decide whether and how it might need to be transformed. Based on the global model, is range size associated with form of migration? How much of the variance in your measure of range size is associated with Migration behavior style?
```{r}
#make a histogram to present range size
hist(d2$Range.Size)
```
```{r}
d2 <- d2 %>% mutate(logRS = log(Range.Size)) #make new variable in dataset
```
```{r}
class(d2$Migration) #check Migration classification status
d2 <- d2 %>% mutate(Migration = as.factor(Migration))
#ANOVA
geo_mig <- aov(data = d2, logRS ~ Migration)
summary(geo_mig)
```
**Question:** Based on the global model, is range size associated with form of migration?\
**Answer:** Yes\

**Question:** How much of the variance in your measure of range size is associated with Migration behavior style?\
**Answer:** I am not sure what you mean by this question?

• Given the regression coefficients returned in output of the model, which Migration categor(ies) are different than the reference level? What level is the reference level? Relevel and assess differences among the remaining pair of Migration categories. 
```{r}
ref_1 <- d2 %>% mutate(Migration = relevel(Migration, ref = 1))
relevel_r1 <- aov(data = d2, logRS ~ Migration)
summary(relevel_r1)

ref_2 <- d2 %>% mutate(Migration = relevel(Migration, ref = 2))
relevel_r2 <- aov(data = d2, logRS ~ Migration)
summary(relevel_r2)

ref_3 <- d2 %>% mutate(Migration = relevel(Migration, ref = 3))
relevel_r3 <- aov(data = d2, logRS ~ Migration)
summary(relevel_r3)
```
**Question:** Given the regression coefficients returned in output of the model, which Migration categor(ies) are different than the reference level?\
**Answer:** None of them appear to be different.\ 

**Question:** What level is the reference level?\
**Answer:** Aren't they all  the reference level?\

• Also conduct a post-hoc Tukey Honest Significant Differences test to also evaluate which Migration categories differ “significantly” from one another (see Module 20).
```{r}
geo_mig <- aov(data = d2, logRS ~ Migration )
posthoc2 <- TukeyHSD(geo_mig, which = "Migration", ordered = TRUE, conf.level = 0.95)
posthoc2 
plot(posthoc2, xlim = c(0, 4))
```

• Winnow your original data to just consider birds from the Infraorder “Passeriformes” (song birds).
```{r}
passer <- d2 %>% filter(Order1 == "Passeriformes")
```

• Run separate one-factor ANOVA analyses to look at the association between [1] relative beak length and Primary.Lifestyle and between [2] relative beak length and Trophic.Level. In doing so…

**ANOVA analysis**
```{r}
#relative beak length to Primary.Lifestyle
beak_len_pl <- aov(data = passer, rel_beak_len ~ Primary.Lifestyle) 
summary(beak_len_pl)
```
```{r}
#relative beak length to Trophic.Level
beak_len_tl <- aov(data = passer, rel_beak_len ~ Trophic.Level)
summary(beak_len_tl)
```

• Make boxplots of response variable by each predictor and by the combination of predictors.
```{r}
#Boxplot of Primary.Lifestyle
beak_prime <- ggplot(d2, aes(x = rel_beak_len, y = Primary.Lifestyle )) + geom_boxplot()
beak_prime
```

```{r}
#Boxplot of Trophic.Level
beak_troph <- ggplot(d2, aes(x = rel_beak_len, y = Trophic.Level)) + geom_boxplot()
beak_troph
```

• Run linear models for each predictor separately and interpret the model output.
```{r}
#linear model for relative beak length with passer PL & TL
beak_len_pl_pass <- lm(data = passer, rel_beak_len ~ Primary.Lifestyle) 
summary(beak_len_pl_pass)

beak_len_tl_pass <- lm(data = passer, rel_beak_len ~ Trophic.Level) 
summary(beak_len_tl_pass)
```

• Run a two-factor model to look at the association between relative beak length and both Primary.Lifestyle and Trophic.Level among the passariforms.
```{r}
#run the two-factor model with ANOVA
tfm <- aov(data = passer, rel_beak_len ~ Primary.Lifestyle + Trophic.Level)
summary(tfm)
```
**Question:** Based on the model output, what would you conclude about how relative beak length is related to these two variables?\
**Answer:** They all appear to be significantly associated with each other.\

• Finally, run an additional two-way model with the same dataset and predictors, but adding the possibility of an interaction term. To do this, you should modify your model formula using the colon operator (:) to specify the interaction, e.g., relative beak length ~ Primary.Lifestyle + Trophic.Level + Primary.Lifestyle:Trophic.Level.
```{r}
#Use the colon operator(:) at the end of the two-factor ANOVA 
tfm2 <- aov(data = passer, rel_beak_len ~ Primary.Lifestyle + Trophic.Level + Primary.Lifestyle:Trophic.Level)
summary(tfm2)
```
**Question:** Based on the model output, what would you now conclude about how relative beak length is related to these two variables?\
**Answer:** Even as they intersect, the relative beak lengths are still significantly associated. 

• Use the interaction.plot() function to visualize the interaction between Primary.Lifestyle and Trophic.Level (see Module 20).
```{r}
#interaction plot of PL and TL
interaction.plot(
  x.factor = passer$Primary.Lifestyle,
  xlab = "Primary Lifestyle",
  trace.factor = passer$Trophic.Level,
  trace.label = "Trophic.Level",
  response = passer$rel_beak_len,
  fun = base::mean,
  ylab = "Relative Beak Length"
)
```